---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-",
  message = FALSE, warning = FALSE, error = FALSE, tidy = TRUE
)
library(epubr)
```

# epubr

[![CRAN status](http://www.r-pkg.org/badges/version/epubr)](https://cran.r-project.org/package=epubr)
[![CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/epubr)](https://cran.r-project.org/package=epubr) 
[![Rdoc](http://www.rdocumentation.org/badges/version/epubr)](http://www.rdocumentation.org/packages/epubr)
[![Travis-CI Build Status](https://travis-ci.org/leonawicz/epubr.svg?branch=master)](https://travis-ci.org/leonawicz/epubr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/leonawicz/epubr?branch=master&svg=true)](https://ci.appveyor.com/project/leonawicz/epubr)
[![Coverage Status](https://img.shields.io/codecov/c/github/leonawicz/epubr/master.svg)](https://codecov.io/github/leonawicz/epubr?branch=master)
[![gitter](https://img.shields.io/badge/GITTER-join%20chat-green.svg)](https://gitter.im/leonawicz/epubr)

## Read EPUB files in R

Read metadata and textual content of epub files. 

`epubr` provides functions supporting the reading and parsing of internal e-book content from EPUB files. E-book metadata and text content are parsed separately and joined together in a tidy, nested tibble data frame. 

E-book formatting is non-standard enough across all literature that no function can curate parsed e-book content across an arbitrary collection of e-books, in completely general form, resulting in a singular, consistently formatted output containing all the same variables.

EPUB file parsing functionality in this package is intended for relatively general application to arbitrary EPUB e-books. However, poorly formatted e-books or e-books with highly uncommon formatting may not work at all with this package.
Text is read 'as is'. Additional text cleaning should be performed by the user at their discretion, such as with functions from packages like `tm` or `qdap`.

This package provides sufficient core functionality at this time. Subsequent versions will include a more robust set of features.

## Installation

Install the development version from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("leonawicz/epubr")
```

## Example

```{r ex}
file <- system.file("dracula.epub", package = "epubr")
(x <- epub(file))

x$data[[1]]
```

```{r is_md, echo = FALSE}
is_md <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github-ascii_identifiers"
```

```{r md_pkg_ref, echo = FALSE, results = "asis", eval = is_md}
cat('## Reference\n\n[Complete package reference and function documentation](https://leonawicz.github.io/epubr/)')
```
